No need to know: astrophysics-agnostic spectral siren cosmology

### 2024-02-14 src/tex/ms.tex
[showyourwork] overleaf sync

\subsection{Option 1: include parametric models}
In this Section, we first show that fitting an incorrect functional form to the mass distribution of CBCs biases the inference of cosmological parameters when using the spectral siren methodology.
We then demonstrate that our flexible model alleviates this bias without the need to know the mass distribution's morphology \emph{a priori}.
To illustrate this, we generate a catalog of \variable{output/num_found_events.txt}GW events from an underlying mass distribution described by the \plp{} model, with hyperparameters consistent with those found in \citet{o3b_pop}.
This simulated catalog is designed to be representative of the gravitational wave transient catalog after the LVK's fifth observing run.
The details of the data simulation are described in Appendix~\ref{ap:data generation}.
We then use this catalog to infer \Ho{} using three different models for the source-frame mass distribution:
\begin{enumerate}
    \item the \plp{} model, which includes the true mass distribution within its hyperprior
    \item a broken power law, which does not include the true mass distribution within its hyperprior, and 
    \item the flexible, Gaussian process-based model described in Section~\ref{sec:model}, which can closely approximate the morphology of the true mass distribution, along with several other morphologies.
\end{enumerate}
The results of each of these fits are shown in Figure~\ref{fig:O5_GP}. 
The \result{left} panel shows the inferred source frame mass distribution for each of the considered models, and the \result{right} panel shows the corresponding posteriors on \Ho{}.

@@

We have repeated this analysis with \result{50} separate simulated catalogs, and found that the power law model causes us to infer a \result{higher/lower} value of \Ho{} at greater than \result{X}-$\sigma$, \result{Y}\% of the time, whereas the Gaussian process-based model reaches that level of bias only \result{Z}\% of the time, and the \plp{} model reaches the same level of bias \result{A}\% of the time.
This indicates that prior knowledge of the shape of the mass distribution is not required to perform a spectral siren measurement, so long as strong assumptions about the shape of the mass distribution are not made.

@@

The resulting inferred mass distribution is shown in Figure~\ref{fig:O5_GP}, along with the corresponding posterior on \Ho.
The inferred mass distribution closely resembles the true, simulated distribution, and the injected value of \Ho (\result{67.66 km/s/Mpc}) is consistent with its inferred posterior distribution (\result{X}).
Additionally, the posterior on \Ho is distinct from its prior distribution, indicating that the data is informative despite the flexibility of the population model.

>>>>

\subsection{Option 1: include parametric models}
In this Section, we first show that fitting an incorrect functional form to the mass distribution of CBCs biases the inference of cosmological parameters when using the spectral siren methodology.
We then demonstrate that our flexible model alleviates this bias without the need to know the mass distribution's morphology \emph{a priori}.
To illustrate this, we generate a catalog of \variable{output/num_found_events.txt}GW events from an underlying mass distribution described by the \plp{} model, with hyperparameters consistent with those found in \citet{o3b_pop}.
This simulated catalog is designed to be representative of the gravitational wave transient catalog after the LVK's fifth observing run.
The details of the data simulation are described in Appendix~\ref{ap:data generation}.
We then use this catalog to infer \Ho{} using three different models for the source-frame mass distribution:
\begin{enumerate}
    \item the \plp{} model, which includes the true mass distribution within its hyperprior
    \item a broken power law, which does not include the true mass distribution within its hyperprior, and 
    \item the flexible, Gaussian process-based model described in Section~\ref{sec:model}, which can closely approximate the morphology of the true mass distribution, along with several other morphologies.
\end{enumerate}
The results of each of these fits are shown in Figure~\ref{fig:O5_GP}. 
The left panel shows the inferred source frame mass distribution for each of the considered models, and the right panel shows the corresponding posteriors on \Ho{}.

@@

We have repeated this analysis with \result{50} separate simulated catalogs, and found that the broken power law model causes us to infer a \result{higher/lower} value of \Ho{} at greater than \result{X}-$\sigma$, \result{Y}\% of the time, whereas the Gaussian process-based model reaches that level of bias only \result{Z}\% of the time, and the \plp{} model reaches the same level of bias \result{A}\% of the time.
This indicates that prior knowledge of the shape of the mass distribution is not required to perform a spectral siren measurement, so long as strong assumptions about the shape of the mass distribution are not made.

@@

The resulting inferred mass distribution is shown in Figure~\ref{fig:O5_GP}, along with the corresponding posterior on \Ho.
The inferred mass distribution closely resembles the true, simulated distribution, and the injected value of \Ho (\variable km/s/Mpc) is consistent with its inferred posterior distribution (\result{X}).
Additionally, the posterior on \Ho is distinct from its prior distribution, indicating that the data is informative despite the flexibility of the population model.
